The transformer architecture introduced by Vaswani et al. represents a fundamental departure from recurrent processing. Multi-head self-attention mechanisms enable direct pairwise interactions between all sequence positions, eliminating the sequential bottleneck inherent in recurrent architectures. Position-invariant attention combined with learned positional encodings provides both parallelizability and the capacity to model long-range dependencies.

This era witnessed the critical integration of reinforcement learning techniques with large-scale language modeling. Schulman et al.'s Proximal Policy Optimization algorithm provided a stable and sample-efficient method for policy gradient optimization, balancing exploration with exploitation through clipped surrogate objectives. Christiano et al. demonstrated that human preferences over trajectory pairs could serve as reward signals for reinforcement learning, eliminating the need for hand-crafted reward functions. This work established reinforcement learning from human feedback as a paradigm for aligning model behavior with human values.

The shift from task-specific architectures to general-purpose pretraining emerged as a dominant paradigm. ELMo demonstrated that contextualized word representations learned from language modeling transfer effectively to downstream tasks. BERT's bidirectional pretraining through masked language modeling and GPT's autoregressive approach established complementary methodologies for learning universal text representations. These models revealed that massive self-supervised pretraining on unlabeled text captures linguistic structure, world knowledge, and reasoning capabilities transferable across diverse applications.

\textbf{Papers in this section:}
\begin{itemize}
    \item \textbf{Vaswani et al. (2017)}: Attention Is All You Need---introduced transformer architecture with multi-head self-attention.
    \item \textbf{Schulman et al. (2017)}: Proximal Policy Optimization---stable policy gradient algorithm with clipped objectives.
    \item \textbf{Christiano et al. (2017)}: Deep RL from Human Preferences---reinforcement learning from preference comparisons without reward engineering.
    \item \textbf{Peters et al. (2018)}: ELMo---contextualized word representations from bidirectional language models.
    \item \textbf{Devlin et al. (2018)}: BERT---bidirectional pretraining via masked language modeling.
    \item \textbf{Radford et al. (2018)}: GPT-1---generative pretraining with autoregressive language modeling.
    \item \textbf{Raffel et al. (2019)}: T5---unified text-to-text transformer framework.
\end{itemize}
